---
title: "Access Open Data with the Junar API"
author: "Frans van Dunné"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction
The Junar API is the basis for a number of Open Data initiatives in Latin
America and the USA. The `junr` package is a wrapper to make it easier to
access data made public through the Junar API. Some examples of implementations
are: [the City of Pasadena](http://data.cityofpasadena.net/home/), and [the
City of San Jose](http://data.sanjoseca.gov/home/). Others are listed on the
[Junar website](http://junar.com/).

## Usage
As an example we will use the data from the [Costa Rican President's
Office](http://datosabiertos.presidencia.go.cr/home).The first step is to
access the website offering the open data to identify the base URL and to
obtain an API Key to get access to the Junar API that hosts the data. You will
find both on the [developers page of the Open Data Costa Rica
site](http://datosabiertos.presidencia.go.cr/developers/). 

Below we use a test API Key so that all the examples will run. You may want to
get your own API Key instead to run the examples below. Note that with Junar
each URL has its own API key.

```{r}
library(junr)
base_url <- "http://api.datosabiertos.presidencia.go.cr/api/v2/datastreams/"
api_key <- "0bd55e858409eefabc629b28b2e7916361ef20ff" 
```

Now that we have the basic information for a connection we can quickly check
what data is available behind this URL.

```{r, eval=FALSE}
get_index(base_url, api_key)
```

The `get_index`function returns the complete list of available data with all
meta-data included as a data frame. 

To get only a list of the global unique identifiers (GUID) of the data sets,
you can use `list_guid`. 


```{r}
list_guid(base_url, api_key)
```

This has the benefit that a reference can be made to the GUID based on position
number. For example:

```{r}
pres_list <-list_guid(base_url, api_key)
pres_list[3]
```

And the same numbers can be used based on the full title with `list_titles`. 

```{r}
list_titles(base_url, api_key)
```

Both `list_guid` and `list_titles` where set up for convenience only because
the results tend to fit in the console window making it easier to read.

Obviously, if you have the GUID of the data that interest you, you can use this
directly to make a call to read all the data. For example, to view all public
purchasing of the presidential office in Costa Rica:

```{r}
data_guid <- "COMPR-PUBLI-DEL-MINIS"
purchasing_data <- get_data(base_url, api_key, data_guid)
```

With `View(purchasing_data)` you can check whether the data have been
downloaded correctly, and have a quick visual check on the mode of the data
(see below to convert currency data from text to numeric).

### Determine data dimensions
On data platforms that run Junar, many data sets are just tables of data that
has already been analyzed and summarized. It is not immediately obvious which
sets contain many data points, and which sets contain only a few rows.

The function `get_dimensions` will download all data sets offered through the
base URL and determine how many rows and columns are available in each one. It
is useful to make a quick assessment of the data available. However, please
note that it may take a while before the function finishes, especially if there
are many GUID's.


```{r}
get_dimensions(base_url, api_key)
```

### Clean up currency data
In the example data above, and possibly in more Junar implementations, we need
to clean up any data related to currency values. In our case we need to found
all currency symbols (Costa Rica Colon) and all the comma's separating
thousands. As they stand these values are text strings, and cannot be converted
directly to numeric without removing the symbols and commas.

There are two utilities to help cleaning the currency data: `clean_currency`
and `get_currency_symbol`.  For example:

```{r}
currency_data <- get_data(base_url, api_key, "LICIT-ADJUD-POR-LOS-MINIS")
currency_data$`Monto Adjudicado` <- clean_currency(currency_data$`Monto Adjudicado`)  
```
